---
title: "Aphanomyces spp. analysis for publication"
author: "Sidhant Chaudhary"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: united
    code_folding: hide
    toc: yes
    toc_float: yes
    fig_caption: yes
    number_sections: yes
    df_print: tibble
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
                      warning=F,
                      message = F,
                      fig.align='center', out.width='100%', fig.asp=1)


# R setup template
rm(list = ls())

# packages and libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, tidytext, janitor, openxlsx, conflicted, skimr, Hmisc, knitr, 
               lme4, lmerTest,
               lsmeans, emmeans, multcomp, multcompView,
               ggpubr, ggcorrplot, ggpubr, ggrepel, GGally, viridis) 

# tidymodels_prefer()

# conflicts: identical function names from different packages
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("lmer", "lmerTest")

# ggplot theme
theme_set(theme_test() + 
          theme(
            #axis.title = element_text(size = 14),
            axis.text = element_text(color = "black"),
            strip.background = element_rect(fill = "skyblue"),
            strip.text = element_text(hjust = 0),
            # legend.text = element_text(size = 12),
            # legend.title = element_text(size = 14), 
            legend.position = "bottom"
          ))
```

# Aim, Hypothesis, and objectives

-   to investigate the distribution of the three different Aphanomyces species: A. euteiches, A. cochlioides, and A. cladogamus in southern Sweden, and to some extent in Denmark and Lithuania.

    -   (the hypothesis is that the presence of different Aphanomyces spp. follow the cultivation of the host plants regionally, but that A. cladogamus is found on a wider geographic scale due to its broad host range.)

-   different species of Aphanomyces co-exist, due to their similar ecological niche.

-   to test if pathogen presence is associated with soil chemical properties, especially Ca.

    -   Moreover, to test whether high calcium levels correlate with low disease pressure.

# Data

```{r}
dta <- read.xlsx("Biotest lista Sc.xlsx",sheet= "Aph. Index + soil analysis 1 MW") %>% as_tibble()
dta <- 
  dta %>%
  clean_names()

colnames(dta)

pea <- 
  dta %>%
  select(id, region, contains("soil"), contains("pea")) %>%
  pivot_longer(cols = contains("index"), names_to = "rep", values_to = "disease") %>%
  mutate(crop = "Pea", 
         rep = str_replace_all(rep, "[^0-9]", "")
         ) %>%
  mutate_if(is.character, as.factor) %>%
  rename(pathogen = note_pea)

sugar_beet <- 
  dta %>%
  select(id, region, contains("soil"), contains("sugar_beet")) %>%
  pivot_longer(cols = contains("index"), names_to = "rep", values_to = "disease") %>%
  mutate(crop = "Sugar beet", 
         rep = str_replace_all(rep, "[^0-9]", "")
         ) %>%
  mutate_if(is.character, as.factor) %>%
  rename(pathogen = note_sugar_beet)

spinach <- 
  dta %>%
  select(id, region, contains("soil"), contains("spinach")) %>% 
  pivot_longer(cols = contains("index"), names_to = "rep", values_to = "disease") %>%
  mutate(crop = "Spinach", 
         rep = str_replace_all(rep, "[^0-9]", "")
         ) %>%
  mutate_if(is.character, as.factor) %>%
  rename(pathogen = note_spinach)

dta_long <- 
  bind_rows(pea, sugar_beet, spinach) %>%
  mutate(pathogen = as.character(pathogen)) %>%
  mutate(pathogen = str_replace(pathogen, "A\\.", "Aphanomyces")) %>%
  mutate(pathogen = factor(pathogen),
         pathogen = fct_recode(pathogen, "None" = "No pathogen found"),
         pathogen = fct_relevel(pathogen, 
                                "Aphanomyces cladogamus",
                                "Aphanomyces cochlioides",
                                "Aphanomyces euteiches",
                                "Aphanomyces cladogamus + Rhizoctonia",
                                "Aphanomyces cochlioides + Rhizoctonia",
                                "Aphanomyces euteiches + Phytophthora",
                                "Aphanomyces sp. + Pythium",
                                "Pythium",
                                "Pythium + Nematodes",
                                "Rhizoctonia",
                                "Rhizoctonia + Pythium",
                                "Rhizoctonia + Pythium + Nematodes",
                                "Nematodes",
                                "None"
                                )) %>%
  relocate(id, region, crop, pathogen, rep, disease)


```

# Exploratory data analysis

# sample distribution

```{r}
library("rnaturalearth")
install.packages("rnaturalearth")
install.packages("regions")
install.packages("eurostat")
install.packages("rnaturalearthdata")
library(regions)
library(eurostat)
sweden <- ne_states(country = "sweden", returnclass = "sf")
map_nuts_2 <- get_eurostat_geospatial(
  resolution = "60", nuts_level = "4",
  year = 2016
)

map_nuts_2 %>%
  filter(CNTR_CODE %in% c("SE", "DK", "LT")) %>% view()
#  as_tibble() %>%
  ggplot() +
  geom_sf()
```


```{r}
dta_long %>%
  group_by(region, id) %>%
  count() %>%
  group_by(region) %>%
  count()
```

# map data
```{r}
# install.packages("maps")
mapdata <- map_data("world")
mapdata %>% 
  as_tibble() %>%
  filter(region %in% c("Sweden", "Denmark", "Lithuania")) %>% view()
  ggplot(aes(x = long, y = lat, group = group, fill = subregion)) + 
  geom_polygon(color = "black")
```


# pathogen distribution each region
```{r}
dta_long %>%
  ggplot(aes(x = region, fill = pathogen)) + 
  geom_bar(position = "fill", color = "black") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```




```{r}
dta_long %>%
  group_by(region, pathogen, crop, id) %>%
  summarise(disease_rep = mean(disease, na.rm = T)) %>%
  ggplot(aes(x = pathogen, y = disease_rep)) + 
  geom_hline(yintercept = 0, lty = 2, color = "gray") + 
  geom_jitter(shape = 21, color = "steelblue") +
  geom_boxplot(color = "black", fill = "steelblue", alpha = 0.8, outlier.shape = NA) + 
  geom_text(data=. %>% 
              group_by(crop, pathogen) %>% 
              summarise(n=n()), 
            aes(x=pathogen, y=-3, label= paste0("(n = ", n, ")")), 
            hjust = 1, color= "gray50") +
  stat_summary(fun = "mean", geom = "point", shape = 18, size = 4) + 
  scale_y_continuous(breaks = seq(0,120, by =20), limits = c(-25, 100)) + 
  coord_flip() +
  facet_wrap(~crop) + 
  labs(x = "", y = "Disease (%)")

```

As the paper will only focus on Aphanomyces, I think we should drop the pathogen levels where we found disease development but didn't detect **Aphanomyces** spp. morphologically.

This leaves us with cases where **Aphanomyces** spp. was detected but alongside other pathogens. I'd say that because we do not know how much pathogens other than **Aphnomyces** spp. contributed to the disease, those cases should be skipped for the further analysis as well. In total, they are 6 cases as below:

```{r}
dta_long %>%
filter( pathogen %in% c("Aphanomyces cladogamus + Rhizoctonia",
                                "Aphanomyces cochlioides + Rhizoctonia",
                                "Aphanomyces euteiches + Phytophthora",
                                "Aphanomyces sp. + Pythium")) %>%
  group_by(region, crop, pathogen) %>%
  summarise(disease_rep = mean(disease, na.rm = T))
```

The dataset with **Aphanomyces** spp. and no pathogen level is as follows:

```{r}
dta_aphano <- 
  dta_long %>%
  filter(pathogen %in% c("Aphanomyces cladogamus",
                         "Aphanomyces cochlioides",
                         "Aphanomyces euteiches",
                         "None")) %>%
  mutate(pathogen = fct_drop(pathogen),
         log_soil_ca = log10(soil_ca), 
         log_soil_p = log10(soil_p),
         log_soil_k = log10(soil_k),
         log_soil_mg = log10(soil_mg)) 

glimpse(dta_aphano)

dta_aphano %>%
  group_by(region, pathogen, crop, id) %>%
  summarise(disease_rep = mean(disease, na.rm = T)) %>%
  ggplot(aes(x = pathogen, y = disease_rep)) + 
  geom_hline(yintercept = 0, lty = 2, color = "gray") + 
  geom_jitter(shape = 21, color = "steelblue") +
  geom_boxplot(color = "black", fill = "steelblue", alpha = 0.8, outlier.shape = NA) + 
  geom_text(data=. %>% 
              group_by(crop, pathogen) %>% 
              summarise(n=n()), 
            aes(x=pathogen, y=-3, label= paste0("(n = ", n, ")")), 
            hjust = 1, color= "gray50") +
  stat_summary(fun = "mean", geom = "point", shape = 18, size = 4) + 
  scale_y_continuous(breaks = seq(0,120, by =20), limits = c(-25, 100)) + 
  coord_flip() +
  facet_wrap(~crop) + 
  labs(x = "", y = "Disease (%)")
```

#  ds
```{r}
dta_aphano %>%
  ggplot(aes(x = region, fill = pathogen)) + 
  geom_bar(position = "fill", color = "black") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```


All visualisations until this point are for exploring.

# Data analysis

```{r}

model_full <- lm(disease ~ region*crop*pathogen, data= dta_aphano)

model_full$call
anova(model_full) 
print("Coefficients: (110 not defined because of singularities) \n
      therefore, 3 way interactions are not kept in the model")

# Post hoc tests
lsmeans <- emmeans(model_full, ~ region + crop + pathogen)
lsmeans_df <- as_data_frame(lsmeans)
lsmeans_tidy <- lsmeans %>% tidy()
cld <- cld(lsmeans, Letters = letters)
cld$.group <- gsub(" ", "", cld$.group)

posthoc_full <- left_join(lsmeans_df, lsmeans_tidy, by = c("region", "crop", "pathogen", "df")) %>%
  left_join(., cld, by = c("region", "crop", "pathogen", "df", "emmean", "upper.CL", "lower.CL", "SE"))
posthoc_full

```

```{r}
# Visualisations
posthoc_full %>%
  # mutate(combo = paste0(region, " - ", crop, " - ", pathogen),
  #        combo = fct_reorder(combo, estimate))  %>%
  filter(!is.na(estimate)) %>%
  ggplot(aes(x = estimate, y = fct_reorder(region, estimate), fill = pathogen)) +
  geom_vline(xintercept = 0, lty = 2, color = "gray") + 
  #geom_errorbar(aes(xmin = estimate -std.error, xmax = estimate + std.error), width = 0)  +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0, position = position_dodge(width = 0.6))  +
  scale_fill_viridis( discrete = T, option = "turbo", direction = 1, begin = 0.1, end = 0.9) +
  geom_point(shape = 21, size = 4, position = position_dodge(width = 0.6)) + 
  # geom_text(aes(x = upper.CL + 2, label = .group), position = position_dodge(width = 0.6), 
  #           hjust = 0, vjust = 0.5 ) +
  geom_text(aes(x = upper.CL + 2 , 
                label = ifelse(p.value < 0.001, "***", 
                               ifelse(p.value < 0.01, "**", 
                                      ifelse(p.value < 0.05, "*", "")))
                ),
            vjust = 0.7, hjust = 0, position = position_dodge(width = 0.6)
            ) +
  scale_x_continuous(breaks = seq(0,120, by =20), limits = c(-25, 100)) + 
  facet_grid(~crop, scales = "free_y", space = "free_y") + 
  labs(title = "Disease severity index in 3 crops in 13 regions by different Aphanomyces sp.",
       x = "Disease %", 
       y = NULL,
       fill = NULL, 
       caption = "Points represent the model mean estimate, error bars represent 95 % confidence interval \nEstimates that are significantly different from 0 are represented by * at p < 0.05, ** at p < 0.01, and *** at p < 0.001")

```

## Distribution of different *Aphanomyces* spp. 

## Co-existence of different *Aphanomyces* spp.

## Pathogen presence association with soil chemical properties, especially Ca

```{r}
## foodhills data
foodhills <- read.xlsx("Foodhills statistik 29 augusti 2023.xlsx") %>% as_tibble()

foodhills <- foodhills %>%
  mutate(pathogen = as.factor(pathogen),
         log_ca = log10(Ca)) %>%
  mutate(pathogen = fct_recode(pathogen, "None" = "No pathogen found")) %>%
  mutate(aphano_binary = as.numeric(ifelse(str_detect(pathogen, "Aphano"), 1, 0)) )
```


```{r}
dta_aphano %>% view()
  filter(rep == 1) %>%
  mutate(aphano_binary = ifelse(pathogen == "None", "None", "Aphanomyces spp.")) %>%
  #count(aphano_binary)
  pivot_longer(soil_p_h:log_soil_mg) %>%
  ggplot(aes(x = value, y = pathogen)) +
    geom_boxplot(outlier.shape = NA) + 
    geom_jitter(aes(fill = region), shape = 21) +
    facet_wrap(~name, scales = "free_x")
  
foodhills %>%
  pivot_longer(c(Ca, log_ca)) %>%
  ggplot(aes(x = value, fill = factor(aphano_binary), group = aphano_binary)) + 
  # geom_boxplot(ou tlier.shape = NA) +
  geom_histogram(position = "identity", color = "black", alpha = 0.7)  + 
  facet_grid(aphano_binary~name, scales = "free")

foodhills %>%
  count(aphano_binary) %>%
  mutate(prop = n/sum(n))

shapiro.test(foodhills$log_ca[foodhills$log_ca < 1.8])
hist(foodhills$log_ca, col = "lightblue", main = "Histogram")
qqnorm(foodhills$Ca); qqline(foodhills$Ca, col = 2)
ggqqplot(foodhills$log_ca[foodhills$log_ca < 1.8], shape = 21)
```

fitting a binomial model

```{r}
model <- glm(aphano_binary ~ log10(Ca), 
             family = binomial(), 
             data = foodhills)
summary(model)
exp(coef(model))

```


```{r}
plot(model)
coef <- coef(model)
odds_ratios <- exp(coef(model))
conf_intervals <- exp(confint(model))

coef[1]

rownames(conf_intervals)

# results table
result_table <- 
  as.data.frame(
    cbind(coef = rownames(coef), 
          odds_ratios, 
          conf_intervals) 
  ) %>% 
  clean_names()
result_table
result_table %>% 
  gather(key = "Interval", value = "value", -coef) %>%
  as_tibble() %>%
  mutate(value = as.numeric(value), 
         value_back_trans = 10^value)

cat("For a one-unit increase in the log10_ca, the odds of the response being 1 decrease by approximately", round(1/0.20549670568165, 2), "times")

cat("For a 10 fold increase in the ca, the odds of the response being 1 decrease by approximately", round(1/0.20549670568165, 2), "times")

cat("For a one-unit increase in the log10_ca, the odds of the response being 1 decrease by approximately", round((1/0.20549670568165 - 1)*100, 2), "%")


cat("For a 10 fold increase in the ca, the odds of the response being apanomyces decrease by approximately", round((1/0.20549670568165 - 1)*100, 2), "%")

calculate_prob_change <- function(prob_base, odds_ratio) {
  prob_change <- (odds_ratio * prob_base) / (1 + prob_base * (odds_ratio - 1))
  return(prob_change)
}
prob_change <- 0.20549670568165/(1 + 0.20549670568165)
prob_change
```


```{r}
# setting limits for x axis
x_limits <- challenger %>%
  dplyr::summarise(min = 0, max = max(temp)+10)

# creating artificial obs for curve smoothing -- several points between the limits
x <- seq(x_limits[[1]], x_limits[[2]], by=0.5)

# artificial points prediction
# see: https://stackoverflow.com/questions/26694931/how-to-plot-logit-and-probit-in-ggplot2
temp.data = data.frame(temp = x) #column name must be equal to the variable name

# Predict the fitted values given the model and hypothetical data
predicted.data <- as.data.frame(
  predict(model_fit, 
          newdata = temp.data, 
          type="link", se=TRUE)
  )

# Combine the hypothetical data and predicted values
new.data <- cbind(temp.data, predicted.data)
##### --------------------------------------------------------------------------

# Compute confidence intervals
std <- qnorm(0.95 / 2 + 0.5)
new.data$ymin <- model_fit$family$linkinv(new.data$fit - std * new.data$se)
new.data$ymax <- model_fit$family$linkinv(new.data$fit + std * new.data$se)
new.data$fit <- model_fit$family$linkinv(new.data$fit)  # Rescale to 0-1

p <- ggplot(challenger, aes(x=temp, y=resp))+ 
  geom_point(colour = plotly_palette[1])+ 
  geom_ribbon(data=new.data, 
              aes(y=fit, ymin=ymin, ymax=ymax), 
              alpha = 0.5, 
              fill = '#FFF0F5')+
  geom_line(data=new.data, aes(y=fit)) + 
  labs(x="Temperature", y="Estimated Fail Probability")+
  ggtitle("Predicted Probabilities for fail/orings with 95% Confidence Interval")+
  theme_bw()+
  theme(panel.border = element_blank(), plot.title = element_text(hjust=0.5))

p

```



```{r}
# putting it in our dataframe
foodhills_predicted <- 
  foodhills %>%
  mutate(log_odds_aphano_binary = coef[1] + coef[2]*log10(Ca),
         odds_aphano_binary = exp(log_odds_aphano_binary),
         odds_aphano_binary_inv = 1/odds_aphano_binary,
        prob = odds_aphano_binary/(1+odds_aphano_binary)) 

foodhills_predicted %>%
  ggplot(aes(x = Ca, y = prob)) +
  geom_line() + 
  #scale_x_log10() + 
  scale_y_continuous(limits = c(0,1))


```

When the odds value is smaller than 1, interpretation is a little bit more complicated. When this is the case, we have to take the inverse value (i.e. 1 divided by the odds) to facilitate interpretation.

The interpretation is then how LESS likely it is to observe the event of interest. For water content, the odds is 0.984. The inverse is:
1/0.984 = 1.0159.

This means that a one-unit increase in water content decreases the likelihood of observing Galumna sp. by 1.0159. We can also subtract 1 from the odds value to obtain a percentage:
( 1.0159 − 1 ) ∗ 100 = 1.59  %.


```{r}
foodhills$.pred <- model$fitted.values
foodhills %>% 
  ggplot(aes(y = .pred, x = Ca, color = factor(aphano_binary))) + 
  geom_jitter(shape = 21, width = 100) 


foodhills %>%
  ggplot(aes(x = Ca, y= aphano_binary)) + 
  scale_x_log10() +
  geom_jitter(aes(fill = factor(aphano_binary)), alpha = 0.8,
              size = 2,
              color = "black", shape = 21, height = .05) + 
  stat_smooth(method = "glm", method.args = list(family=binomial), se = TRUE) 

```

## Are high calcium levels correlate with low disease pressure?

# Conclusions

# Session information

```{r}
sessionInfo()
```
